01 - TreehopperFitness: Quality control

### 2023-08-24 scripts/Publilia-01-qc.Rmd

## `Publilia-02-strats.Rmd`
```{r}
write_rds(All_data_v6, here("saved_rds/All_data_v6.rds"))
write_rds(clutch_status_guarding_summary_v2, here("saved_rds/clutch_status_guarding_summary_v2.rds"))
write_rds(last_day_fID, here("saved_rds/last_day_fID.rds"))
```

@@

```{r}
strats_data <- left_join(clutch_status_guarding_summary_v2, last_day_fID)

visited_1st_4d_data <- strats_data %>%
  group_by(location) %>%
    mutate(visited_1st_4d = any(!mom_at_init & as.numeric(visit_starts-initiation_date)<=4)) %>%
  ungroup() %>%
  mutate(`visited in 4 days` = ifelse(visited_1st_4d | num_at_init>1,
                                      "Visited", "Not visited"),
         ants = ifelse(ants_removed, "Ants removed", "Control"),
         est_hatch = as_date(ifelse(is.na(hatch_date),
                                 initiation_date + 20,
                                 hatch_date)),
         est_rel2h = visit_ends-est_hatch)
```

@@

# 9.2 - Threshold for number of days ants were on the plant
Should I be more lenient with my threshold, so I can have more even sample sizes between "ants" and "no ants"?
- ants_ever = n_days_ants>0: F = 83, T = 143
- antsP_gt2d = n_days_ants>2: F = 123, T = 103 * most even sample size
- antsP_gt3d = n_days_ants>3: F = 133, T = 90
```{r}
ant_n_days <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  mutate(ant_treatment=ifelse(ants_removed,"Ants removed","Control"),
         Dorm=substr(location,1,1)) %>%
  group_by(Date, location) %>%
    summarize(any_antsP_day = any(ants_P>0, na.rm=T),
              females_removed = first(females_removed),
              ant_treatment = first(ant_treatment),
              Dorm = first(Dorm)) %>%
  ungroup() %>%
  group_by(location) %>%
    summarize(n_days_ants = sum(as.numeric(any_antsP_day)),
              females_removed = first(females_removed),
              ant_treatment = first(ant_treatment),
              Dorm = first(Dorm))%>%
  ungroup() %>%
  mutate(antsP_gt2d = !is.na(n_days_ants)&n_days_ants>2,
         antsP_gt3d = !is.na(n_days_ants)&n_days_ants>3,
         ants_everP = is.na(n_days_ants)|n_days_ants>0)

@@

@@

>>>>

@@

```{r}
visited_1st_4d_data <- left_join(clutch_status_guarding_summary_v2, last_day_fID) %>%
  group_by(location) %>%
    mutate(visited_1st_4d = any(!mom_at_init & as.numeric(visit_starts-initiation_date)<=4)) %>%
  ungroup() %>%
  mutate(`visited in 4 days` = ifelse(visited_1st_4d | num_at_init>1,
                                      "Visited", "Not visited"),
         ants = ifelse(ants_removed, "Ants removed", "Control"),
         est_hatch = as_date(ifelse(is.na(hatch_date),
                                 initiation_date + 20,
                                 hatch_date)),
         est_rel2h = visit_ends-est_hatch)
```

@@

## 9.2 - Threshold for number of days ants were on the plant
Should I be more lenient with my threshold, so I can have more even sample sizes between "ants" and "no ants"?
- ants_ever = n_days_ants>0: F = 83, T = 143
- antsP_gt2d = n_days_ants>2: F = 123, T = 103 * most even sample size
- antsP_gt3d = n_days_ants>3: F = 133, T = 90
```{r}
ant_n_days <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  mutate(ant_treatment=ifelse(ants_removed,"Ants removed","Control"),
         Dorm=substr(location,1,1)) %>%
  group_by(Date, location) %>%
    summarize(any_antsP_day = any(ants_P>0, na.rm=T),
              females_removed = first(females_removed),
              ant_treatment = first(ant_treatment),
              Dorm = first(Dorm)) %>%
  ungroup() %>%
  group_by(location) %>%
    summarize(n_days_ants = sum(as.numeric(any_antsP_day)),
              females_removed = first(females_removed),
              ant_treatment = first(ant_treatment),
              Dorm = first(Dorm))%>%
  ungroup() %>%
  mutate(antsP_gt2d = !is.na(n_days_ants)&n_days_ants>2,
         antsP_gt3d = !is.na(n_days_ants)&n_days_ants>3,
         ants_everP = is.na(n_days_ants)|n_days_ants>0)

@@

## 9.3 - Ant arrival times

```{r}
antP_arrival_afh_data <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch,
           is_antsP_afh=ants_P&afh) %>%
    filter(is_antsP_afh) %>%
    slice_head(n=1) %>% 
  mutate(antP_arrival_afh=Date-est_hatch) %>%
  select(location, antP_arrival_afh, ants_removed, females_removed,
         initiation_date, est_hatch) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

antP_last_b4h_data <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch,
           is_antsP_bfh=ants_P&b4h) %>%
    filter(is_antsP_bfh) %>%
    slice_tail(n=1) %>% 
  mutate(antP_last_b4h=Date-est_hatch) %>%
  select(location, antP_last_b4h, ants_removed, females_removed,
         initiation_date, est_hatch) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

antL_arrival_afh_data <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch,
           is_antsL_afh=ants_L&afh) %>%
    filter(is_antsL_afh) %>%
    slice_head(n=1) %>% 
  mutate(antL_arrival_afh=Date-est_hatch) %>%
  select(location, antL_arrival_afh, ants_removed, females_removed,
         initiation_date, est_hatch) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

antL_last_b4h_data <- All_data_v6 %>%
  left_join(select(visited_1st_4d_data, location, initiation_date, est_hatch, ants_removed, females_removed)) %>%
  filter(!is.na(initiation_date)) %>%
  distinct(Date,location,ants_removed,females_removed,initiation_date,est_hatch,ants_L,ants_P) %>%
  group_by(location) %>%
    mutate(b4h=Date>=initiation_date&Date<est_hatch,
           afh=Date>=est_hatch,
           is_antsL_bfh=ants_L&b4h) %>%
    filter(is_antsL_bfh) %>%
    slice_tail(n=1) %>% 
  mutate(antL_last_b4h=Date-est_hatch) %>%
  select(location, antL_last_b4h, ants_removed, females_removed,
         initiation_date, est_hatch) %>%
  ungroup() %>%
  mutate(Dorm=substr(location,1,1),
         ants_treatment=ifelse(ants_removed,"Ants removed","Control"))

ant_data <- ant_counts %>%
  left_join(antP_last_b4h_data) %>%
  left_join(antP_arrival_afh_data) %>%
  left_join(antL_last_b4h_data) %>%
  left_join(antL_arrival_afh_data)
```

@@

## `Publilia-02-strats.Rmd`
```{r}
strats_data <- visited_1st_4d_data_tb  %>%
  # Add ant arrival data
  left_join(ant_data) %>%
  # Add day of first visit bins
  left_join(first_visit_cuts) %>%
  mutate(vis_cat = ifelse(num_at_init > 1,
                          "0",
                          ifelse(is.na(first_visit_starts),
                                 "never visited",
                                 ifelse(first_visit_cut=="(0,4]", "1 to 4", "5+"))),
         vis_cat = factor(vis_cat, levels = c("0","1 to 4",
                                              "5+","never visited"))) %>%
  # Accommodate different tie-breaking procedures (wrt choosing a single female as initiator if multiple present at hatch)
  # 1) Longest initiator
  mutate(initL = init_first_dur_rank == 1) %>% 
  # 2) Shortest initiator
  #  i) break ties by picking marked females over unmarked
  group_by(location, init_first_dur_shortest) %>% 
    mutate(initS_tmp = ifelse(n()>1, !grepl("Uf", Female_ID), TRUE)) %>% 
  ungroup() %>%
  #  ii) finally picking at random (only one case, identical predictor vars)
  group_by(location, init_first_dur_shortest, initS_tmp) %>%
    mutate(initS = init_first_dur_shortest & initS_tmp & row_number()==1) %>%
  ungroup()

## `Publilia-02-strats.Rmd`
```{r}
write_rds(All_data_v6, here("saved_rds/All_data_v6.rds"))
# write_rds(clutch_status_guarding_summary_v2, here("saved_rds/clutch_status_guarding_summary_v2.rds"))
# write_rds(last_day_fID, here("saved_rds/last_day_fID.rds"))
write_rds(strats_data, here("saved_rds/strats_data.rds"))
```
